An Efficient Calculus for Generalised Parikh Images of Regular Languages

### 2022-07-06 main.tex
Examples

@@

  Note that \Expand{} is taken directly from the existing formulation of
  \cref{eq:generate-parikh}.
\end{proof}

@@

We return again to our example in \cref{sec:introduction:parikh}, where we
compute the whole Parikh image of the product of $\AcaOrBc{}$ and
$\SomethingCSomething$ under the constraint that there are more instances of
letters a than b. We omit the definition of the map $\Map$ that maps each
transition to the corresponding increment vector (e.g.
\Map($\FromLabelTo{S}{a}{A}) = \begin{bmatrix}1\\0\\0\end{bmatrix}$),
since it should be apparent by now.

In the interest of space, we use a similar technique as in
\cref{fig:example:single:equivalences} to combine the application of $\FlowEq$
with the definition of $\Filter$.

  We define $\Filter$ after substitutions from both initial applications of
  $\FlowEq$ as follows, note the element-wise expansion for the $\Sigma$ labels:

% Propagated equations
    \begin{equation*}
      \begin{aligned}
        & \Filter(\FromLabelTo{S}{a}{A}) & = 1 - \TransitionVar_6 \\
        & \Filter(\FromLabelTo{S}{b}{B}) & = \TransitionVar_4 + \TransitionVar_6 \\
        & \Filter(\FromLabelTo{A}{c}{A})  & = \TransitionVar_3  \\
        & \Filter(\FromLabelTo{B}{b}{S}) & = 2\TransitionVar_6 + \TransitionVar_4 \\
        & \Filter(\FromLabelTo{A}{a}{F}) & = 1 - \TransitionVar_6 \\
        & \Filter(\FromLabelTo{B}{c}{F}) & = \TransitionVar_6 \\
        % Other automaton
        & \Filter(\FromLabelTo{S}{\Sigma}{S}) & = \TransitionVar_{7a}, \TransitionVar_{7b}, \TransitionVar_{7c} \\
        & \Filter(\FromLabelTo{S}{c}{F}) & = 1 \\
        & \Filter(\FromLabelTo{F}{\Sigma}{F}) & = \TransitionVar_{9a}, \TransitionVar_{9b}, \TransitionVar_{9c} \\
      \end{aligned}
    \end{equation*}

\begin{prooftree}
  \infer0[\PresburgerClose{}]{
    \begin{matrix}
      1 = \TransitionVar_{10} \\
      \TransitionVar_3 = \TransitionVar_{11} \\
      1 = \TransitionVar_{12} \\
      1 = \TransitionVar_{11} \\
      2 > 0
    \end{matrix}  
  }
  \infer1[\Subsume{}, \Expand{}, \Subsume{}]{
    \begin{matrix}
      1 = \TransitionVar_{10} \\
      \TransitionVar_3 = \TransitionVar_{11} \\
      1 = \TransitionVar_{12} \\
      1 = \TransitionVar_{11} \\
      \Image{}_{\Automaton', \Map}(\Filter, 
      \begin{bmatrix}
        2 \\
        0 \\
        1
        \end{bmatrix}) \land \\
      \Connected(\SomethingCSomething{}, \Filter) \land \\
        2 > 0
    \end{matrix}  
  }
  \infer1[Eq. reasoning]{
    \begin{matrix}
      1 = \TransitionVar_{10} \\
      \TransitionVar_3 = \TransitionVar_{11} \\
      1 = \TransitionVar_{12} \\
      1 = \TransitionVar_{11} \\
      \Image{}_{\Automaton', \Map}(\Filter, 
      \begin{bmatrix}
        a \\
        0 \\
        c
        \end{bmatrix}) \land \\
      \TransitionVar_6 = 0 \land \\
      \TransitionVar_4 = 0 \land \\
      2\TransitionVar_6 + \TransitionVar_4 = 0 \land \\
      \begin{bmatrix}
        a \\
        b \\
        c
        \end{bmatrix} = \begin{bmatrix}
          2  \\
          0 \\
          1
        \end{bmatrix} \land
        \\
        \begin{bmatrix}
          a \\
          b \\
          c
          \end{bmatrix} = \begin{bmatrix}
            \TransitionVar_{7a} + \TransitionVar_{9a} \\
            0 \\
            \TransitionVar_{7c} + \TransitionVar_{9c} + 1
          \end{bmatrix} \land \\
      \Connected(\SomethingCSomething{}, \Filter) \land \\
        a > 0
    \end{matrix}  
  }
\end{prooftree}

\begin{prooftree}
  \infer0[\Materialise]{
    \begin{matrix}
      \TransitionVar_6 = 0 \\
      \TransitionVar_4 = 0 \\
      2\TransitionVar_6 + \TransitionVar_4 = 0 \\
      \begin{bmatrix}
        a \\
        b \\
        c
        \end{bmatrix} = \begin{bmatrix}
          2  \\
          0 \\
          \TransitionVar_3 
        \end{bmatrix}
        \\
        \begin{bmatrix}
          a \\
          b \\
          c
          \end{bmatrix} = \begin{bmatrix}
            \TransitionVar_{7a} + \TransitionVar_{9a} \\
            0 \\
            \TransitionVar_{7c} + \TransitionVar_{9c} + 1
          \end{bmatrix} \\
      \Connected(\SomethingCSomething{}, \Filter) \land \\
      \Image{}_{\AcaOrBc{}\times\SomethingCSomething{}, \Map}(\Filter, 
      \begin{bmatrix}
        a \\
        0 \\
        c
        \end{bmatrix}) \land \\
        a > 0
    \end{matrix}  
  }
  \infer1[\Subsume{}]{
  \begin{matrix}
    \TransitionVar_6 + \TransitionVar_4 = 0 \\
    1 - \TransitionVar_6 > 0 \\
    \TransitionVar_6 = 0 \\
    \TransitionVar_4 = 0 \\
    \TransitionVar_3 > 0 \\
    2\TransitionVar_6 + \TransitionVar_4 = 0 \\
    \begin{bmatrix}
      a \\
      b \\
      c
      \end{bmatrix} = \begin{bmatrix}
        2  \\
        0 \\
        \TransitionVar_3 
      \end{bmatrix}
      \\
      \begin{bmatrix} % What to do about sigma!?
        a \\
        b \\
        c
        \end{bmatrix} = \begin{bmatrix}
          \TransitionVar_{7a} + \TransitionVar_{9a} \\
          0 \\
          \TransitionVar_{7c} + \TransitionVar_{9c} + 1
        \end{bmatrix} \\
    \Connected(\AcaOrBc{}, \Filter) \land \\
    \Connected(\SomethingCSomething{}, \Filter) \land \\
    \Image{}_{\AcaOrBc{}\times\SomethingCSomething{}, \Map}(\Filter, 
    \begin{bmatrix}
      a \\
      0 \\
      c
      \end{bmatrix}) \land \\
      a > 0
  \end{matrix}
  }
  \infer1[Eq. reasoning]{
    \begin{matrix}
      \begin{bmatrix}
        a \\
        b \\
        c
        \end{bmatrix} = \begin{bmatrix}
          2 - 2\TransitionVar_6 \\
          2\TransitionVar_4 + 3\TransitionVar_6 \\
          \TransitionVar_3 + \TransitionVar_6
        \end{bmatrix}
        \\
        \begin{bmatrix} % What to do about sigma!?
          a \\
          b \\
          c
          \end{bmatrix} = \begin{bmatrix}
            \TransitionVar_{7a} + \TransitionVar_{9a} \\
            \TransitionVar_{7b} + \TransitionVar_{9b} \\
            \TransitionVar_{7c} + \TransitionVar_{9c} + 1
          \end{bmatrix} \\
      \Connected(\AcaOrBc{}, \Filter) \land \\
      \Connected(\SomethingCSomething{}, \Filter) \land \\
      \Image{}_{\AcaOrBc{}\times\SomethingCSomething{}, \Map}(\Filter, 
      \begin{bmatrix}
        a \\
        b \\
        c
        \end{bmatrix}) \land \\
        a > b
    \end{matrix}
  }
  \infer1[\ExpandM]{\Image{}_{\AcaOrBc{}\times\SomethingCSomething{}, \Map}(\Filter, \begin{bmatrix}
    a \\
    b \\
    c
    \end{bmatrix}) \land a > b}
\end{prooftree}

>>>>

\Fudge{I don't understand.}
  % \[
  % \begin{aligned}
  %   & \inferrule
  %   {\Connected(\Automaton, \Filter) \land \FlowEq(\Automaton, \Filter) \land \MonoidElement = \sum_{\Transition \in \Transitions_\Automaton} \Filter(\Transition) \cdot \Map(\Transition), \SomeInequalities, \SomeClause}
  %   {\SinglePredicateInstance, \SomeInequalities, \SomeClause} &\beta \models R_0 \iff \exists i : \beta \models R_i
  % \end{aligned}
  % \]

@@

  Note that \Expand{} is taken directly from the existing formulation of
  \eqref{eq:generate-parikh}.
\end{proof}

@@

We return again to our example in \cref{sec:introduction:parikh}, where we
compute the whole Parikh image of the product of $\AcaOrBc{}$ and
$\SomethingCSomething$ under the constraint that there are more instances of
letters a than b. This time our monoid $\Monoid$ is 3-dimensional vectors with
element-wise addition, and $\Map$ that maps each transition to the corresponding
increment vector, e.g $\Map(\FromLabelTo{S}{a}{A}) = \VectorLiteral{1,0,0}$.

In the interest of space, we refer back to
\cref{fig:example:single:equivalences} for the definitions of $\Filter$ for
$\AcaOrBc$ and only define $\Filter$ for $\SomethingCSomething{}$ after
substitutions as follows. Note the expansion of the $\Sigma$ labels:
    \begin{equation*}
      \begin{aligned}
        % & \Filter(\FromLabelTo{S}{a}{A}) & = 1 - \TransitionVar_6 \\
        % & \Filter(\FromLabelTo{S}{b}{B}) & = \TransitionVar_4 + \TransitionVar_6 \\
        % & \Filter(\FromLabelTo{A}{c}{A})  & = \TransitionVar_3  \\
        % & \Filter(\FromLabelTo{B}{b}{S}) & = 2\TransitionVar_6 + \TransitionVar_4 \\
        % & \Filter(\FromLabelTo{A}{a}{F}) & = 1 - \TransitionVar_6 \\
        % & \Filter(\FromLabelTo{B}{c}{F}) & = \TransitionVar_6 \\
        % Other automaton
        & \Filter(\FromLabelTo{S}{\Sigma}{S}) & = \VectorLiteral{\TransitionVar_{7a}, \TransitionVar_{7b}, \TransitionVar_{7c}} \\
        & \Filter(\FromLabelTo{S}{c}{F}) & = 1 \\
        & \Filter(\FromLabelTo{F}{\Sigma}{F}) & = \VectorLiteral{\TransitionVar_{9a}, \TransitionVar_{9b}, \TransitionVar_{9c}} \\
      \end{aligned}
    \end{equation*}

\begin{prooftree}
  \infer0[\PresburgerClose{}]{
    \begin{matrix}
      1 = \TransitionVar_{10} \land 
      \TransitionVar_3 = \TransitionVar_{11} \land \\
      1 = \TransitionVar_{12} \land 
      1 = \TransitionVar_{11} \land 
      2 > 0
    \end{matrix}  
  \infer1[\Subsume{}, \Expand{}, \Subsume{}]{
    \begin{matrix}
      1 = \TransitionVar_{10} \land 
      \TransitionVar_3 = \TransitionVar_{11} \land  \\
      1 = \TransitionVar_{12} \land 
      1 = \TransitionVar_{11} \land \\
      \Image{}_{\Automaton', \Map}(\Filter, 
      \VectorLiteral{2, 0, 1}) \land 
      \Connected(\SomethingCSomething{}, \Filter) \land
        2 > 0
    \end{matrix}  
  \infer1[\EquationReasoning{}]{
    \begin{matrix}
      1 = \TransitionVar_{10} \land
      \TransitionVar_3 = \TransitionVar_{11} 
      1 = \TransitionVar_{12} \land
      1 = \TransitionVar_{11} \\
      \Image{}_{\Automaton', \Map}(\Filter, 
      \VectorLiteral{
        \TransitionVar_{7a} + \TransitionVar_{9a},
        0,
        \TransitionVar_{7c} + \TransitionVar_{9c} + 1}) \land \\
        \TransitionVar_6 + \TransitionVar_4 = 0 \land 
        1 - \TransitionVar_6 > 0 \land
        \TransitionVar_6 = 0 \land 
        \TransitionVar_4 = 0 \land
        \TransitionVar_3 > 0 \land
        2\TransitionVar_6 + \TransitionVar_4 = 0 \land \\
          \VectorLiteral{
        \TransitionVar_{7a} + \TransitionVar_{9a},
        0,
        \TransitionVar_{7c} + \TransitionVar_{9c} + 1}
        = \VectorLiteral{
          2,
          0,
          \TransitionVar_3} \land \\
      \Connected(\SomethingCSomething{}, \Filter) \land
      \land \TransitionVar_{7a} + \TransitionVar_{9a} > 0
    \end{matrix}  
  \infer1[\Materialise]{
    \begin{matrix}
      \TransitionVar_6 + \TransitionVar_4 = 0 \land 
      1 - \TransitionVar_6 > 0 \land
      \TransitionVar_6 = 0 \land 
      \TransitionVar_4 = 0 \land
      \TransitionVar_3 > 0 \land
      2\TransitionVar_6 + \TransitionVar_4 = 0 \land \\
      \VectorLiteral{
        \TransitionVar_{7a} + \TransitionVar_{9a},
        0,
        \TransitionVar_{7c} + \TransitionVar_{9c} + 1}
        = \VectorLiteral{
          2,
          0,
          \TransitionVar_3} \land \\
      \Connected(\SomethingCSomething{}, \Filter) \land \\
      \Image{}_{\AcaOrBc{}\times\SomethingCSomething{}, \Map}(\Filter, 
      \VectorLiteral{\TransitionVar_{7a} + \TransitionVar_{9a}, 0, \TransitionVar_{7c} + \TransitionVar_{9c} + 1}) \land \TransitionVar_{7a} + \TransitionVar_{9a} > 0
    \end{matrix}
  \infer1[\Subsume{}]{
  \begin{matrix}
    \TransitionVar_6 + \TransitionVar_4 = 0 \land 
    1 - \TransitionVar_6 > 0 \land
    \TransitionVar_6 = 0 \land 
    \TransitionVar_4 = 0 \land
    \TransitionVar_3 > 0 \land
    2\TransitionVar_6 + \TransitionVar_4 = 0 \land \\
    \VectorLiteral{
      \TransitionVar_{7a} + \TransitionVar_{9a},
      0,
      \TransitionVar_{7c} + \TransitionVar_{9c} + 1}
      = \VectorLiteral{
        2,
        0,
        \TransitionVar_3} \land \\
    \Connected(\AcaOrBc{}, \Filter) \land 
    \Connected(\SomethingCSomething{}, \Filter) \land \\
    \Image{}_{\AcaOrBc{}\times\SomethingCSomething{}, \Map}(\Filter, 
    \VectorLiteral{\TransitionVar_{7a} + \TransitionVar_{9a}, 0, \TransitionVar_{7c} + \TransitionVar_{9c} + 1}) \land \TransitionVar_{7a} + \TransitionVar_{9a} > 0
  \end{matrix}
  \infer1[\EquationReasoning]{
    \begin{matrix}
      \VectorLiteral{a, b, c} = \VectorLiteral{
          2 - 2\TransitionVar_6,
          2\TransitionVar_4 + 3\TransitionVar_6,
          \TransitionVar_3 + \TransitionVar_6
        } \land
        \\
        \VectorLiteral{a, b, c} = \VectorLiteral{
            \TransitionVar_{7a} + \TransitionVar_{9a},
            \TransitionVar_{7b} + \TransitionVar_{9b},
            \TransitionVar_{7c} + \TransitionVar_{9c} + 1
          } \land \\
      \Connected(\AcaOrBc{}, \Filter) \land 
      \Connected(\SomethingCSomething{}, \Filter) \land \\
      \Image{}_{\AcaOrBc{}\times\SomethingCSomething{}, \Map}(\Filter, 
      \VectorLiteral{a, b, c}) \land a > b
    \end{matrix}
  \infer1[\ExpandM]{\Image{}_{\AcaOrBc{}\times\SomethingCSomething{}, \Map}(\Filter, \VectorLiteral{a, b, c}) \land a > b}
\end{prooftree}
